A Methodology and Implementation of Automated Emissions Harmonization for Use in Integrated Assessment Models

### 2017-11-23 paper.tex
intro edits
finished dvv edits in intro

typo in intro

methods revisions

typo in texit

some final cleaning of meths

text edits in results

added figure for rad forcing

fix acks

updating forcing fig and text

clearing up some results text

rm wm cmd
so it begins

%%% ELS includes
% * <sfujimori112256@gmail.com> 2017-05-23T14:15:19.404Z:
% 
% comment from SF
% 
% ^ <sfujimori112256@gmail.com> 2017-05-24T18:49:42.852Z.
\documentclass[review]{elsarticle}
\usepackage{lineno,hyperref}
\modulolinenumbers[5]
\journal{Environmental Modelling \& Software}
\bibliographystyle{elsarticle-num}
%%%

@@

\include{intro}
\include{meth}
\include{results}
% * <sfujimori112256@gmail.com> 2017-05-25T05:18:35.016Z:
% 
% I think the current results are basically fine (sorry I have not read in detailed though).  What we should show and discuss are pretty much depending on how we evaluate these results. So, at this stage, I would like to put more comments on this section
% 
% 
% ^.
\include{future}

@@

The authors would like to acknowledge a number of colleagues who helped
contribute both discussion and feedback regarding this work including Drs. Elmar
Kriegler, Gunnar Luderer, and Joeri Rogelj. The authors further wish to thank
the Global Environment Facility for their generous support. This project has
received funding from the European Union’s Horizon 2020 research and innovation
programme under grant agreement No 641816.

>>>>

\documentclass[review]{elsarticle}
\usepackage{lineno,hyperref}
\modulolinenumbers[5]
\journal{Environmental Modelling \& Software}
\bibliographystyle{elsarticle-num}
%%%

@@

\section{Introduction}

Integrated Assessment Models (IAMs) are tools used to understand the complex
interactions between energy, economy, land use, water, and climate systems. IAMs
provide global projections of systemic change by dividing the world into a
number of representative regions (typically 10 to 30), the definition of which
is distinct for each model \cite{krey_global_2014}. Results from IAMs are
integral in a number of international studies, which notably include projections
of climate and economic futures. Recently, the IAM community has developed
scenarios based on the Shared Socio-economic Pathways (SSPs)
\cite{van_vuuren_energy_2017, fricko_marker_2017, fujimori_ssp3:_2017,
  calvin_ssp4:_2017, kriegler_fossil-fueled_2017} which quantify a variety of
potential global futures. The SSPs are designed to be used in research that
include earth system model (ESM) simulations, climate impact, adaptation and
climate mitigation studies \cite{vuuren_new_2013}.

While IAMs are implemented in myriad ways, including simulation and
optimization, the core inputs and outputs are similar across different
models. Modeling teams incorporate data on energy systems, land use, economics,
demographics and emissions sources and concentrations, among other data, in
order to provide consistent existing trajectories of modeled variables. The
models then provide estimates of future trajectories of these variables under
various socio-economic and technological assumptions as well as proposed policy
constraints, e.g., targets for future Greenhouse Gas (GHG) emissions.

The emissions trajectories calculated by IAMs are critical inputs for ongoing,
worldwide scientific community efforts in the Coupled Model Intercomparison
Project Phase 6 (CMIP6) \cite{eyring_overview_2016}, which is utilizing a number
of marker SSP scenarios developed by the IAM community (ScenarioMIP)
\cite{oneill_scenario_2016}. These trajectories are endogenously calculated by
modeling the individual technologies and sectors that contribute towards the
emissions of different air pollutants and GHGs as well as various mitigation
technologies. However, the historical emissions starting points of models can
differ by large amounts depending on the region, sector, and emissions species.

In practice, IAMs calculate the total source intensity of emitting technologies,
for example the total activity of coal power plants in China, and incorporate
emissions-intensity factors for individual gas species, for example the quantity
of sulfur emissions from coal plants per megawatt-hour of production. Models are
generally \textit{calibrated} to historical data sources in one or more base
years. Results in the historical period may differ between models as a result of
the sometimes large uncertainties in historical data sets. Models can also
differ in their choice of base-year, which may lag behind available inventory
data. In addition, models have varying sectoral, regional, and fuel
aggregations.

The global climate change community has recently developed a new global
historical emissions data set for both anthropogenic emissions (i.e., the
Community Emissions Data System (CEDS) \cite{hoesly_historical_2017} and
open-burning land-use and land-use change (LULUC) emissions
\cite{van_marle_historic_2017}) which, in conjunction with the SSP IAM
trajectories, will be used for climate-related modeling exercises of CMIP6.

When participating in intercomparison exercises in which a consistent historical
starting point is required (e.g., in CMIP6), model teams incorporate a single,
common historical data set through \textit{harmonization}. Harmonization
refers to the process of adjusting model results to match a selected
historical time series such that the resulting future trajectories are also
consistent with the original modeled results. In the emissions context, this
means that each individual combination of model region, model sector, and
emissions species must be harmonized. Depending on the total number of model
regions, sectors, and emissions species, this can require the selection of
thousands to tens-of-thousands of harmonization methods.

Harmonization has been addressed in previous studies as it is a common practice
in the IAM and climate change communities. For example,
\cite{meinshausen_rcp_2011} describes the use of scaling routines for the 5
regions used in the Special Report on Emissions Scenarios (SRES)
\cite{nakicenovic2000}; however, only total emissions were harmonized in the
exercise, thus there is no sectoral dimension. Further,
\cite{rogelj_discrepancies_2011} describes the impacts of choosing various
harmonization routines on future trajectories. During the evaluation of the
Representative Concentration Pathways (RCPs), IAM results have been harmonized
by sector and the 5 RCP global regions
\cite{vuuren_representative_2011}. Importantly, the choice of harmonization
method to date has been determined by individual experts and has generally been
applied to all trajectories for a given class of emissions species.

Climate modeling efforts have continued to progress, demanding increased spatial
and sectoral resolution from IAMs. Furthermore, a new generation of climate
scenarios which combines aspects of both the RCPs and SSPs have been developed
in order to incorporate both physical and socio-economic detail. In order to
address the growing dimensionality of model outputs and support ongoing scenario
generation and analysis efforts while still providing a consistent and
scientifically rigorous harmonization procedure, an automated process for
determining harmonization methods is preferred. The use of an automated,
documented, and openly available harmonization mechanism additionally allows for
full procedural reproducibility and for direct participation by additional
modeling teams not involved in the original exercise.

The remainder of this paper describes the methodology and implementation of the
harmonization software \code{aneris} \cite{matthew_gidden_2017_802832}, written
in the Python programming language (detailed documentation available
\href{http://mattgidden.com/aneris/}{online}). Section \ref{sec:meths} provides
a detailed description of the underlying mathematical components of
\code{aneris} as well as the procedural workflow. The results of applying the
automated harmonization mechanism on two example IAM scenarios, one with
emissions growth and another with emissions mitigation, is presented in Section
\ref{sec:results}. Finally, the general effectiveness and potential future
improvements on the automated methodology is discussed in Section
\ref{sec:future}.
\section{Methodology \& Implementation}\label{sec:meths}

\subsection{Harmonization Methods}

IAM emission results are provided along temporal (normally half decade or
decade), spatial (i.e., model regions), gas species, and sectoral
dimensions. Each individual temporal trajectory, i.e., unique combinations of
regions ($r$), species ($g$), and sectors ($s$), must be harmonized to the
initial modeling period. Given a model trajectory, $m_{r, g, s}(t)$, historical
values, $h_{r, g, s}(t)$, and model base year, $t_i$, a harmonized trajectory
needs to be calculated. The \textit{harmonization quality} of a trajectory,
i.e., how well a given harmonization algorithm performs, depends on a number of
factors. Of chief import is the faithful representation of the original,
unharmonized, trajectory as well as the representation of negative trajectories
(i.e., if a trajectory becomes negative, both the timing and total magnitude
should be as close as possible) which are of critical importance for cumulative
\cotwo calculations.

In previous studies \cite{meinshausen_rcp_2011,rogelj_discrepancies_2011}, two
\textit{families} of methods have been used: those that operate on the ratio of
base year values (i.e., $\frac{h(t_i)}{m(t_i)}$) and those that operate on the
offset of base year values (i.e., $h(t_i) - m(t_i)$). A number of the classic
methods are implemented in \code{aneris} including ratio-convergence shown in
Equation \ref{eqs:ratio}, offset-convergence shown in Equation \ref{eqs:offset},
and interpolation shown in Equation \ref{eqs:interpolate}. The convergence
factor, $\beta$, scales linearly from 1 to 0 over $[t_i, t_f)$ and is shown in
  Equation \ref{eqs:factor}. In all equations the region, species, and sector
  indices have been dropped for clarity. Each equation is a function of time,
  model trajectory, historical trajectory, base year ($t_i$), and a convergence
  year ($t_f$), at which point the harmonized trajectory converges to the
  unharmonized trajectory. \codeb{aneris} provides a number of methods to choose
  from for each of the harmonization families. A summary of all available
  methods is provided in Table \ref{tab:meths}.

\begin{equation}\label{eqs:factor}
  \beta(t, t_i, t_f) =
  \begin{cases}
    1 - \frac{t - t_i}{t_f - t_i},& \text{if } t < t_f\\
    0,                        & \text{otherwise}
  \end{cases}
\end{equation}

\begin{equation}\label{eqs:ratio}
  m^{rat}(t, m, h, t_i, t_f) = [\beta(t, t_i, t_f) (\frac{h(t_i)}{m(t_i)} - 1) + 1] m(t)
\end{equation}

\begin{equation}\label{eqs:offset}
  m^{off}(t, m, h, t_i, t_f) = \beta(t, t_i, t_f) (h(t_i) - m(t_i)) + m(t)
\end{equation}

\begin{equation}\label{eqs:interpolate}
  m^{int}(t, m, h, t_i, t_f) =
  \begin{cases}
    \beta(t, t_i, t_f) (h(t_i) - m(t_f)) + m(t_f), & \text{if } t < t_f\\
    m(t), & \text{otherwise}
  \end{cases}
\end{equation}

\begin{table}[h!]
\centering
\caption{All Harmonization Methods Provided in \code{aneris}}
\label{tab:meths}
\begin{tabular}{|l|l|l|}
\hline
Method Name                             & Harmonization Family & Convergence Year\\
\hline
\code{constant\_ratio}                  & ratio              & $t_f = \infty$\\
\code{reduce\_ratio\_<year>}            & ratio              & $t_f = $\code{<year>}\\
\code{constant\_offset}                 & offset             & $t_f = \infty$\\
\code{reduce\_offset\_<year>}           & offset             & $t_f = $\code{<year>}\\
\code{linear\_interpolate\_<year>}      & interpolation      & $t_f = $\code{<year>}\\
\hline
\end{tabular}
\end{table}

\subsection{Default Method Decision Tree}\label{sec:tree}

A \textit{decision tree} approach has been implemented in \codeb{aneris} which
provides a systematic and documented decision-making process to determine the
preferred harmonization algorithm. In order to provide reasonable
\textit{default} methods, the historical trajectory, unharmonized model
trajectory, and relative difference between history and model values in the
harmonization year are analyzed. The decision tree used in this analysis is a
result of collaborative efforts between IAM teams and is shown graphically in
Figure \ref{fig:decision_tree}.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{decision_tree.pdf}
    \caption[]{
      \label{fig:decision_tree}
      The default method decision tree used in the \code{aneris} software
      library. For all decisions, upper (purple) branches represent a ``yes''
      response and lower (orange) branches represent a ``no'' response. The
      coefficient of variation, $c_v$, is defined in Equation \ref{eqs:cov},
      $dH$ is defined as $\left|\frac{h(t_i) - m(t_i)}{h(t_i)}\right|$, and
      decision-making thresholds for $c_v$ and $dH$ are described below. Methods
      labeled in \textit{green} are likely to closely match unharmonized
      results, methods in \textit{yellow} will likely somewhat match
      unharmonized results, and methods in \textit{red} can be expected to have
      a large relative difference between harmonized and unharmonized results.}
  \end{center}
\end{figure}

A number of characteristics impact the decision of which default method to
select based on the effect of the characteristic on the potential harmonized
trajectory. For example, it is possible for models to report zero values in the
harmonization year in situations in which technologies are introduced in future
time periods in regions or for sectors which produce an emissions species that
is absent in the initial modeling period. In such cases, an offset method is
required as a ratio method would mask future emissions and erroneously harmonize
the trajectory.

In most cases, however, models do report values in the harmonization
year. Figure \ref{fig:cases} displays a number of example trajectories which
highlight the possible issues resulting from harmonizing model results in
different contexts. When model and historical values are relatively close, a
convergence method is chosen in order to be as representative as possible to the
underlying unharmonized model results (Figure \ref{fig:cases}, Panel
\code{a}). If values are not close, the constant ratio method is chosen in order
to provide reasonable trajectories that still incorporate modeled effects
(Figure \ref{fig:cases}, Panel \code{b}).

Models can additionally report negative emissions in certain contexts which must
be taken into account during harmonization. Such a case is possible for gas
species which can be extracted from the environment and stored, as is the case
for \cotwo in future scenarios with climate mitigation. If a model provides a
trajectory that transitions from positive to negative emissions and base year
results are similar, then a convergence method is used in order to guarantee
capture of this transition in a representative fashion (Figure \ref{fig:cases},
Panel \code{c}). If the discrepancy in base year results is large, it is
possible for a negative trajectory to be inappropriately harmonized to a
positive, but decreasing, trajectory. As such, the constant ratio method is
chosen (Figure \ref{fig:cases}, Panel \code{d}).

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{cases.pdf}
    \caption[]{
      \label{fig:cases}
      The effect of different harmonization routines on model trajectories under
      ``normal'' circumstances (Panel \code{a}), when there is a large
      difference between historical and model values in the harmonization year
      (Panels \code{b} and \code{d}), and when model trajectories result in
      negative emissions by the end of the modeling time horizon (Panels
      \code{c} and \code{d}). Identical model trajectories are used in each row
      (Panels \code{a}, \code{b}; \code{c}, \code{d}). In each column,
      historical values are increased in the base year by an order of magnitude
      (from 10 to 100). In each Panel, a subset of the potential routines
      provide a better harmonization quality than others as described in the
      text.
    }
  \end{center}
\end{figure}

Temporal variability of the historical trajectory is also an important
characteristic when considering the choice of harmonization method.  Emissions
from forest and grassland fires, for example, vary from year to year due to a
combination of meteorological conditions and anthropogenic drivers. Land use
emissions in many IAMs are modeled using average emission factors and do not
capture conditions in a specific year. A longer convergence horizon is thus
desired in order to incorporate highly variable historical data with modeled
results as is consistency in harmonization method because the effects are
modeled similarly across regions and species. In order to detect emissions with
a high amount of variation, a measure of the coefficient of variation, $c_v$, of
the first derivative of the historical trajectory is calculated using the
standard deviation, $\sigma$, and the mean, $\mu$, as shown in Equation
\ref{eqs:cov}.

The value of $c_v$ is then tested against a threshold, $\tau_{c_v}$. To
determine this threshold, an analysis of the recent CEDS and LUC historical data
has been performed. Figure \ref{fig:cov} shows the distribution of LUC $c_v$s
and non-LUC $c_v$s as determined for historical data aggregated to the model
regions of 5 different IAMs involved in the SSP process. A threshold value of
$\tau_{c_v} = 20$ has been chosen based on these observations as it
optimally divides the two distributions. Importantly, tails of the LUC and
non-LUC overlap, thus there are both false positives (~7\% of non-LUC
trajectories) and false negatives (~10\% of LUC trajectories). However, as any
regional definition is model dependent and thus any regional aggregation is
possible an automated detection mechanism is necessary.

\begin{equation}\label{eqs:cov}
    c_v =  \frac{\sigma(h^{\prime}(t))}{\mu(h^{\prime}(t))}
\end{equation}

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{cov.pdf}
    \caption[]{
      \label{fig:cov}
      The distribution of $c_v$ values for LUC and non-LUC historical
      trajectories is shown. CEDS historical data \cite{hoesly_historical_2017}
      is used for non-LUC data and \cite{van_marle_historic_2017} is used for
      LUC data. All historical data has been aggregated from countries to IAM
      model regional definitions, and all gas species included in the historical
      data sets are included in the analysis. The solid black line indicates the
      threshold value, $\tau_{c_v}$, used by default in \code{aneris}.  }
  \end{center}
\end{figure}

Finally, consideration is taken with respect to the relative difference between
the historic and model values in the harmonization time period. In order to
investigate the possible values that these relative differences can take, the
IAM values used in the SSP and (ongoing) CMIP6 inter-comparison
exercises are used. A distribution of these differences for all models in the
study is presented in Figure \ref{fig:dh}. Given the available data, a threshold
value of $\tau_{dH} = 50$\% was chosen to be used as a default in \code{aneris}.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{dh.pdf}
    \caption[]{
      \label{fig:dh}
      The distribution of relative differences between model and historical
      values in the harmonization year is shown. The solid black line indicates
      the 50\% threshold value, $\tau_{dH}$, used by default in \code{aneris}.
    }
  \end{center}
\end{figure}

\subsection{\code{aneris} Workflow}\label{sec:workflow}

The full \code{aneris} workflow is comprised of a number of components shown
graphically in Figure \ref{fig:workflow}. Unharmonized model data and a
run-control configuration are read in via an Excel spreadsheet. Data is assumed
to be in the IAMC format, i.e., using \code{Model}, \code{Scenario},
\code{Region}, \code{Variable}, and \code{Unit} columns in addition to columns
representing each modeled time period.

Users are able to control the harmonization process via a number of options. The
primary mechanism by which users control the process is by providing
\textit{override} methods for any combination of region and variable (i.e.,
sector and gas species).  In practice, it may be possible that not all default
methods chosen will provide robust harmonized trajectories, especially if there
is a significant difference between historical and model values in the
harmonization year, if there is significant upward or downward movement in the
model trajectory, or if there are known discrepancies in sectoral definition
between the IAM and historical data source. In such cases, override methods can
ameliorate the issues associated with the default method choice. In order to
help identify these cases, harmonization \textit{diagnostics} are provided which
analyze the relative difference between harmonized and unharmonized trajectories
at their mid (if $> 400$\%) and end-points (if $> 200$\%). If override methods
are provided, they are used instead of the default methods as determined by the
default method decision tree. Furthermore, users can set the above-mentioned
thresholds as well as the LUC method used in the decision tree (see Figure
\ref{fig:decision_tree}). Further detail of input parameters can be found
\href{http://mattgidden.com/aneris/config.html}{online}.

\begin{figure}
  \begin{center}
    \includegraphics[height=0.9\textheight]{aneris-workflow.pdf}
    \caption[]{
      \label{fig:workflow}
      The full harmonization process as executed by \code{aneris}. Operations
      that can be configured with user-based input configurations are shown in
      purple. The core harmonization process is shown in yellow.  }
  \end{center}
\end{figure}

Input data then undergoes a cleaning operation, which adds (null) model
trajectories that exist in the historical data set but are not provided by the
model input and detects any issues that would cause the harmonization process to
fail. The methods used to harmonize the data are then determined and the
harmonization process is executed. Upon completion of the harmonization process,
aggregation of common analysis regions is performed. A common regional
aggregation used in the IAM community was defined in the Representative
Concentration Pathways (RCPs) \cite{vuuren_representative_2011}, shown in Figure
\ref{fig:regions}. Finally, any exogenous trajectories the user provides are
added. Exogenous trajectories are normally provided for unmodeled gases with
well-accepted scenario trajectories, e.g., chlorofluorocarbons provided by WMO
\cite{wmo2014}. Upon completion, the harmonized trajectories and meta data
regarding the harmonization process are returned. A description of all returned
meta data is provided in Table \ref{tab:metadata}.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{MESSAGE_11-5regions_map.pdf}
    \caption[]{
      \label{fig:regions}
      The 5 regions used in the RCPs with their 11-region constituents: Asia
      (Central Asia, South Asia, Pacific Asia) [yellows], Latin America
      [magenta], the Middle East and Africa [greens], the OECD (North America,
      Western Europe, and Pacific OECD) [blues], and the Reforming Economies
      (Eastern Europe and Former Soviet Union) [reds].  }
  \end{center}
\end{figure}

\begin{table}[]
\centering
\caption{Meta data provided by the \code{aneris} harmonization routine. This meta data is provided for every combination of region, sector, and emissions species.}
\label{tab:metadata}
\begin{tabular}{|p{2cm}|p{8cm}|}
\hline
\textbf{Column}       & \textbf{Description}                                      \\
\hline
\hline
method       & The harmonization method used.                                               \\
\hline
default      & The default harmonization method as determined by the default decision tree. \\
\hline
override     & The method provided as an override (if any).                                 \\
\hline
offset       & The offset value between history and model in the harmonization year.        \\
\hline
ratio        & The ratio value between history and model in the harmonization year.         \\
\hline
cov          & The coefficient of variation value of the historical trajectory.                           \\
\hline
unharmonized & The unharmonized value in the harmonization year.                            \\
\hline
history      & The historical value in the harmonization year.                             \\
\hline
harmonized   & The resulting harmonized value in the harmonization year.\\
\hline
\end{tabular}
\end{table}
\section{Results}\label{sec:results}

In order to show a representative cross section of the performance of the
\code{aneris} harmonization procedure, we focus on the harmonization of results
of the IAM MESSAGE-GLOBIOM \cite{fricko_marker_2017}. Two scenarios from the SSP
scenario library \cite{Riahi2017153,Rao2017346} are presented. The SSP2, or
``middle of the road'', scenario (referred to as SSP2-Ref) is chosen to be
analyzed because MESSAGE-GLOBIOM is the marker scenario for this SSP. This SSP2
scenario lies between two RCPs, 6 and 8.5, with a radiative forcing level of
approximately 6.5 Wm$^{-2}$. We additionally present the results for the
SSP2-based mitigation scenario leading to a radiative forcing of 4.5 Wm$^{-2}$
(referred to as SSP2-4.5). The SSP2-45 scenario is chosen because mitigation
technologies and policies are enacted causing a general reduction in pollutants
and GHGs, including (eventual) negative CO2 emissions in some regions and
sectors due to carbon capture and sequestration and afforestation. Figure
\ref{fig:kyoto} shows the different trends of Kyoto Gases in each scenarios.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{results_kyoto.pdf}
    \caption[]{
      \label{fig:kyoto}
      Unharmonized Kyoto gas emissions for SSP2-Ref, a scenario with generally
      increasing global emissions trends, and SSP2-45, a scenario with generally
      decreasing global emissions trends.  }
  \end{center}
\end{figure}

MESSAGE-GLOBIOM includes a representation of 11 distinct regions which can be
mapped directly to the 5-region definition used in the RCPs. Historical data is
taken from previously described LUC and anthropogenic sources, which comprise 10
separate pollutant and GHG species and 12 sectors shown in Table \ref{tab:sp}. A
total of 970 distinct trajectories were harmonized for each scenario, and
therefore 1940 trajectories were harmonized in total. \noxx generated from the
Energy sector provides an example of an emissions species and sector in which
all regions were satisfactorily harmonized with the default methods. Figure
\ref{fig:nox} shows the results of harmonization in Asia, and Table
\ref{tab:nox} describes the parameters that underlie the choice of method for
each harmonized trajectory.

\begin{table}[]\begin{minipage}{\textwidth}
  \centering
  \caption{Harmonized Species and Sectors}
  \label{tab:sp}
  \begin{tabular}{|l|l|}
    \hline
    \textbf{Emissions Species}        & \textbf{Sectors}                    \\
    \hline
    \hline
    Black Carbon (BC)                 & Agricultural Waste Burning          \\
    Hexafluoroethane (C\textsubscript{2}F\textsubscript{6})
    \footnote{\label{x_glb_all}
      Global total trajectories are harmonized due to lack of detailed 
      historical data.}
                                      & Agriculture                         \\
    Tetrafluoromethane (CF\textsubscript{4})
    \footnoteref{x_glb_all}               & Aircraft
    \footnote{\label{x_glb}
      Global sectoral trajectories are harmonized due to lack of detailed 
      historical data.}                            \\
    Methane (CH\textsubscript{4})                     & Energy Sector                       \\
    Carbon Dioxide (CO\textsubscript{2})
    \footnote{\label{x_co2}
      A global trajectory for land-use CO\textsubscript{2} is used; non-land-use sectors 
      are harmonized for each model region.}
                                      & Forest Burning                      \\
    Carbon Monoxide (CO)              & Grassland Burning                   \\
    Hydrofluorocarbons (HFCs)
    \footnoteref{x_glb_all}               & Industrial Sector                   \\
    Nitrous Oxide (N\textsubscript{2}O)
    \footnoteref{x_glb_all}
                                      & International Shipping\footnoteref{x_glb}              \\
    Ammonia (\nht)                     & Residential Commercial Other        \\
    Nitrogen Oxides (\nox)             & Solvents Production and Application \\
    Organic Carbon (OC)               & Transportation Sector               \\
    Sulfur Hexafluoride (SF\textsubscript{6})
    \footnoteref{x_glb_all}               & Waste                               \\
    Sulfur Oxides (SO\textsubscript{x})               &                                     \\
    Volatile Organic Compounds (VOCs) &                                     \\
    \hline
  \end{tabular}
\end{minipage}\end{table}

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{example_NOx_Energy_Sector.pdf}
    \caption[]{
      \label{fig:nox}
      \noxx Energy Sector harmonized (solid lines) and unharmonized (dashed lines)
      trajectories for SSP2 and SSP2-45 with historical trajectories (grey
      lines) are presented. The SSP2 reference scenario is shown in Panels
      \textbf{a} and \textbf{c}; the SSP-45 scenario is denoted with ``x''
      markers in Panels \textbf{b} and \textbf{d}. The upper panels show the
      results for endogenously modeled and harmonized regions in Asia while the
      lower panels display the aggregate region results.  
    }
  \end{center}
\end{figure}

\begin{table}[]
\centering
\caption{Key Parameters for Deciding Harmonization Methods for \noxx Emissions in the Energy Sector in Asia}
\label{tab:nox}
\begin{tabular}{|p{1.2cm}|p{.5cm}|p{.5cm}|p{4.5cm}|p{3cm}|}
\hline
\textbf{Region} & \textbf{dH} & $\mathbf{c_v}$ & \textbf{Decision Tree Traversal (Branch and Direction)} & \textbf{Default Method Chosen} \\ \hline
    \hline
CPA             & 0.35        & 2.26          & 1 (no), 2 (no), 3 (yes)                                 & reduce\_ratio\_2080    \\ \hline
PAS             & 0.14        & 1.24          & 1 (no), 2 (no), 3 (yes)                                 & reduce\_ratio\_2080    \\ \hline
SAS             & 0.56        & 0.58          & 1 (no), 2 (no), 3 (no), 4 (no)                          & constant\_ratio        \\ \hline
\end{tabular}
\end{table}

The harmonization of emissions pathways is performed in order to accurately
represent new or updated data sets of historical emissions inventories while also
maintaining consistency with the original, unharmonized pathway. As such, when
the default methods as provided by the harmonization procedure distort or
otherwise sufficiently misrepresent the underlying unharmonized results, an
override method is required to be provided for the trajectory of the region,
sector, and species in question. Of the 970 trajectories, approximately 10\%
were reported as a diagnostic (see Section \ref{sec:workflow}) of which 3.5\%
required the use of harmonization overrides after an initial investigation;
thus, 96.5\% of all trajectories were satisfactorily harmonized using the
default methods. The trajectories that required overrides clustered into two
classifications: regional trajectories whose \textit{magnitude} was overly
distorted and regional trajectories whose \textit{shape} was overly distorted.

Figure \ref{fig:co} presents a case in which the magnitude of a trajectory is
distorted. A large discrepancy ($\sim$300\% relative difference) is observed in
the harmonization year for carbon monoxide (CO) emissions in the industrial
sector specifically for the South Asia (SAS) MESSAGE-GLOBIOM region, which
comprises most of the emissions of the Asian subcontinent. The default method
chosen (\code{constant_ratio}) maintains model trends for the region; however,
overall model results are distorted. By applying a \code{constant_offset}
override, the regional trend and magnitude is maintained. With the new
harmonization method for the SAS region, the global trajectory for industrial CO
also is representative the trends seen in the unharmonized trajectory and the
relative importance of the underlying regional trajectories is maintained.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{results_CO_Industrial_Sector.pdf}
    \caption[]{
      \label{fig:co}
      CO Industrial Sector harmonized and unharmonized emissions are presented
      for SSP2 and SSP2-45 scenarios. Scenarios as denoted identically to Figure
      \ref{fig:nox}. Panels \textbf{a} and \textbf{b} show harmonized and
      overridden-harmonized (respectively) regional trajectories for the 3
      MESSAGE-GLOBIOM regions that comprise the R5ASIA region: Centrally Planned
      Asia (CPA), Other Pacific Asia (PAS), and South Asia (SAS). Notably, the
      SAS regional trajectory displays a distorted trajectory due to the
      harmonization-year difference between history and model results in both
      scenarios. The distortion is large enough to affect global results, as
      shown in Panels \textbf{c} and \textbf{d}.  
}
  \end{center}
\end{figure}

In certain circumstances, the application of the default harmonization methods
can affect not only the magnitude but also the shape of regional
trajectories. Figure \ref{fig:nh3} shows an example case of emissions
trajectories for ammonia (\nht) from the agriculture sector in Asia. Again, the
SAS region shows a large discrepancy in the harmonization year ($>$150\% in this
case). The resulting trajectory harmonized with the default method
(\code{constant_ratio}) provides a large increase after 2080 in the SSP2
reference scenario. Notably, the SSP2-45 scenario is not affected to the same
degree. While this distortion changes the magnitude of the SAS trajectory, it
largely affects the post-2080 shape of the global trajectory (see Figure
\ref{fig:nh3}, panel \textbf{c}). By using a \code{constant_offset} method as an
override, this distortion is addressed and more accurately reflects unharmonized
results in the SAS region, the relative importance between regions, and global
results for agricultural ammonia emissions, each of which contributes to a
better harmonization quality for the harmonized SAS trajectory.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{results_NH3_Agriculture.pdf}
    \caption[]{
      \label{fig:nh3}
      \nhtt agricultural harmonized and unharmonized emissions are presented for
      SSP2 and SSP2-45 scenarios. Scenarios and panel layouts are identical to
      Figure \ref{fig:co}. In this case, the SAS trajectory again shows not only
      a magnitude distortion, but also a shape distortion at the tail of the
      trajectory. Additionally, global trajectories are greatly affected by the
      harmonization method choice (there is $\sim$20\% relative difference
      between trajectories in the reference scenario in 2100). Override methods
      have been applied to correct the distortion.  }
  \end{center}
\end{figure}

We investigate the aggregate effect of harmonization with all necessary override
methods on total anthropogenic radiative forcing projections with the simple
carbon-cycle and climate model, MAGICC6 \cite{meinshausen2011emulating,
  meinshausen2011rcp}, for each harmonized and unharmonized scenario
respectively as shown in Figure \ref{fig:forcing}. We find that the change due
to harmonization is small, ranging between 1 and 2.5\% over the model
horizon. Relative near-term differences persist in the mitigation case (SSP-4.5)
because differences in near-term emissions define to a larger degree the
longer-term forcing outcome due to the cumulative nature of long-lived climate
forcers like \cotwo.  The resulting difference in forcing in 2100 is 0.04
$\frac{\text{W}}{\text{m}^2}$ for SSP2-4.5 and 0.01
$\frac{\text{W}}{\text{m}^2}$ for SSP2-Ref, both of which are well within
acceptable tolerances (e.g., 0.75 $\frac{\text{W}}{\text{m}^2}$ defined for
ScenarioMIP \cite{oneill_scenario_2016}). Thus harmonization is considered to
have a negligible effect on key long-term climate indicators.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{results_forcing.pdf}
    \caption[]{
      \label{fig:forcing}
      The results of the simple climate model, MAGICC6, forced with the SSP2-Ref
      (blue) and SSP2-4.5 (green) harmonized and unharmonized scenarios is
      presented. The radiative forcing trajectories of harmonized and
      unharmonized scenarios are shown in solid lines and dashed lines,
      respectively.  }
  \end{center}
\end{figure}
\section{Discussion \& Future Work}\label{sec:future}

This work presented a novel methodology and Python implementation of automated
emissions harmonization for IAMs, \code{aneris}. An in-depth explanation of the
processes and methods for determining the use of harmonization methods was
provided in Section \ref{sec:meths}. \code{aneris} was able to satisfactorily
harmonized over 96\% of the 1940 individual trajectories that were analyzed in
Section \ref{sec:results}. Of the remaining trajectories, harmonization method
overrides were applied, and example situations in overrides were needed were
discussed.

The automated approach drastically reduces the need for expert opinion in
determining harmonization methods for each individual combination of model
region, sector, and emissions species while still providing a justifiable
explanation for each automated choice of harmonization method based on both the
historical and future emissions trajectories. Furthermore, the automated
approach continues to scale well as models become more detailed in both the
regional and sectoral dimensions. Finally, expert opinion is still allowed to
trump the automated method as determined by the algorithm via method overrides;
however, these cases are clearly documented via the meta data provided as an
output of \code{aneris} and thus can be individually explained. This provides
not only transparency and reproducibility, but also scientific integrity in the
choice of harmonization methods.

The use of an open-source, automated harmonization process also provides
benefits to the wider climate science and IAM communities. By providing a
standard mechanism for harmonization, the IAM community can directly provide
input into the harmonization algorithms and rules for their default
selection. Additionally, modeling teams are easily capable of executing
identical harmonization procedures in order to participate in ongoing and
further iterations of intercomparison exercises and analysis. Future scenario
analyses can also utilize this common harmonization approach such that they are
consistent with prior efforts.

There are a variety of avenues for future improvement of both the \code{aneris}
software and underlying methodology. As with any software project, additional
users will provide use cases for more robust handling of input/output concerns
and corner cases. Further configuration parameters can also be added in the
future in order to provide overrides for all gas species in a given sector or
region. Perhaps the most fruitful investigation will involve further refinement
of the default decision tree introduced in Section \ref{sec:meths}. A key aspect
missing from the decision tree is input from models regarding whether missing
sources or activity levels are the likely cause of a harmonization year
discrepancy (suggesting the use of an offset method) or instead a significant
difference in emissions factors (suggesting the use of a ratio method)
\cite{rogelj_discrepancies_2011}.

This work provides a new direction and framework which the IAM and climate
communities can build upon in order to reduce the necessity of consistent expert
opinion and increase transparency and reproducibility of harmonization
exercises. Furthermore, it provides an open-source, tested, and documented
software library which can be used and improved upon by these communities. Both
of these are clear steps in a positive direction for future climate and
integrated assessment modeling exercises.

@@

The authors would like to acknowledge a number of colleagues who helped
contribute both discussion and feedback regarding this work including Drs. Elmar
Kriegler, Gunnar Luderer, and Joeri Rogelj. This project has received funding
from the European Union’s Horizon 2020 research and innovation programme under
grant agreement No 641816. The authors further wish to thank the Global
Environment Facility for their generous financial support.
